---
title: Coursera-Developing Data Products - Course Project - Shiny Application and Reproducible Pitch
author: "Conrad Frank"
date: "October 11, 2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

 - This presentation is the second part of the assignment of week 4, ***Developing Data Products course from Coursera*** (<https://www.coursera.org/learn/data-products>). 
 
 - The presentation is meant to show information about the first part of the above mentioned assignment (a development project). 

 - The presentation was generated using [***RStudio Presentor***](<https://www.rstudio.com>) framework. 

## Application

 - The first part of the assignment is to develop a web application. The application is named ***GDP forecast*** and is hosted at <https://conradfrank.shinyapps.io/GDP_forecast/>.
 

 - This application shows an example of making a web application using R and [Shiny](https://shiny.rstudio.com) framework together.
 
 - In the application, there are 2 parts. First we select a country and plot the **GDP per capita** (current US$) . The data is derived from **World Bank** <font size="3"> © 2018 The World Bank Group, All Rights Reserved.</font> using the library(**WDI**). This is a Time Series data.
 
 - Second part in the app is to forecast the GDP data for the Country selected from the years 2018 to 2028 which is decided by the user.
For this, library(**forecast**) is used. The [*auto.arima()*](https://www.rdocumentation.org/packages/forecast/versions/8.4/topics/auto.arima) function is used which is a Fit Best ARIMA (AutoRegressive Integrated Moving Average) Model To Univariate Time Series. Then the *forecast()* function is used to derive the predicted GDP for user defined years and plotted.
 
 - The application and the current presentation source codes is uploaded at <https://github.com/conradfrank/shiny_app>. The application comprises 2 files: *ui.R* (UI), *server.R* (backend).
  
## Dataset

Dataset used by the application is the [The World Bank - Databank](http://http://databank.worldbank.org/data/home.aspx) (using *'WDI'* package). The GDP data used is **GDP per capita (current US$)** which has a metadata tag -[*NY.GDP.PCAP.CD*](https://data.worldbank.org/indicator/NY.GDP.PCAP.CD)

The dataset structure: 
```{r data, echo = TRUE,message=FALSE}
library(WDI)
data <- WDI(indicator = "NY.GDP.PCAP.CD", country = 'USA', start = 1990, end = 2017 ) #USA selected for reference only
str(data)
```

## Prediction

Using the *auto.arima()* function, the best fit model of [ARIMA](https://en.wikipedia.org/wiki/Autoregressive_integrated_moving_average) is selected. The [forecast](https://www.rdocumentation.org/packages/forecast/versions/8.4/topics/forecast) function is used to predict the Time series for the number of years (ie from 2018 to 2028) decided by the user.

A plot is generated has consists of the original GDP data from 1990-2017, the fittee values generated from forecast 1990-2017 and the extension plot of the predicted values of GDP. This also plots the C.I. of 80% and 95% 


```{r lib,echo = FALSE, eval=FALSE}
suppressWarnings(suppressMessages(library(WDI)))
suppressWarnings(suppressMessages(library(forecast)))

```

```{r code,echo = TRUE, eval=FALSE}
data <- WDI(indicator = "NY.GDP.PCAP.CD", country = "XYZ", start = 1990, end = 2018 )
# XYZ - is02c format code of the country selected by the user

ARIMAfit = auto.arima(df, approximation=FALSE,trace=FALSE)
#this codes selects the best ARIMA model based on the GDP data

forcst <- forecast(ARIMAfit,h=X)
#X is calculated based on the number of years from 2018 onwards decided by the user.
#The function generates the forecast TS data.
```
